An accelerometer-derived ballistocardiogram method for detecting heartrates in free-ranging marine mammals

### 2022-04-28 analysis/paper/paper.Rmd
Knit figures to PDF

```{r setup, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  cache = TRUE,
  comment = "#>",
  fig.path = "../figures/",
  fig.height = 6,
  fig.width = 6,
  dpi = 300
)
library(cetaceanbcg)
library(tidyverse)
```

@@

While the BCG method presented here holds the potential to mine existing and future marine mammal bio-logging datasets for information about cardiovascular function, it has several limitations compared to ECG methods. Most importantly, BCGs are highly sensitive to movement artifacts [@inanBallistocardiographySeismocardiographyReview2015], so only motionless periods are valid for analysis. This limits the behavioral and physiological contexts in which heartrate may be measured. For example, the BCG is probably an inappropriate method for quantifying the magnitude of surface tachycardia [@goldbogenExtremeBradycardiaTachycardia2019] and exercise modulation of bradycardia [@noren2012], due to movement artifacts during those activities. Secondarily, we did not test whether the BCG is robust to tag placement location. The blue whale data presented in this study was collected when a dorsally-deployed tag slipped to the lateral chest cavity behind a flipper, where it is reasonable to expect greater accelerations caused by heart beats than from a tag farther from the animal's center of mass. It is possible that the ballistic forces generated by heart beats are strong enough to produce an interpretable BCG for a variety of potential tag deployment locations, but this likely varies with animal body size, as well as accelerometer sampling rate and sensitivity.

>>>>

```{r setup, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  cache = TRUE,
  comment = "#>",
  fig.path = "../figures/",
  fig.height = 6,
  fig.width = 6,
  dev = "pdf",
  dpi = 300
)
library(cetaceanbcg)
library(tidyverse)
```

@@

While the BCG method presented here holds the potential to mine existing and future marine mammal bio-logging datasets for information about cardiovascular function, it has several limitations compared to ECG methods. Most importantly, BCGs are highly sensitive to movement artifacts [@inanBallistocardiographySeismocardiographyRemovview2015], so only motionless periods are valid for analysis. This limits the behavioral and physiological contexts in which heartrate may be measured. For example, the BCG is probably an inappropriate method for quantifying the magnitude of surface tachycardia [@goldbogenExtremeBradycardiaTachycardia2019] and exercise modulation of bradycardia [@noren2012], due to movement artifacts during those activities. Secondarily, we did not test whether the BCG is robust to tag placement location. The blue whale data presented in this study was collected when a dorsally-deployed tag slipped to the lateral chest cavity behind a flipper, where it is reasonable to expect greater accelerations caused by heart beats than from a tag farther from the animal's center of mass. It is possible that the ballistic forces generated by heart beats are strong enough to produce an interpretable BCG for a variety of potential tag deployment locations, but this likely varies with animal body size, as well as accelerometer sampling rate and sensitivity.
